---
title: "Replication of Sainburg 2020 and Combination with Smith & Lewicki 2006
"
author: "Nathan Tansey"
date: "2024-10-13"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


<!-- Replication reports should all use this template to standardize reporting across projects.  These reports will be public supplementary materials that accompany the summary report(s) of the aggregate results. -->

## Introduction

During Tim Sainburg's time at UCSD in Tim Gentner's lab, he published several papers, creating and inspiring many of the analysis techniques used to this day within animal vocalization research. Among these is the method described in "Finding, visualizing, and quantifying latent structure across diverse animal vocal repertoires", which is agnostic to specific animal vocalizations. Part of my prior research has been in adaptations of vocalizations to environmental distortion, this toolbox could allow easier comparisons of animals dealing with specific environments. Additionally, understanding and replicating the computational methods will establish another tool to use at baseline for any analysis of ongoing experimentation within the Gentner lab. Due to my computational background, and discussions with Tim Gentner, I will also add the base framework of sparse coding representations as outlined by Smith & Lewicki's "Efficient Auditory Coding", which I have used to success in past research to isolate the essential features of audio.
The first step needed will be data preparation. Due to the high availability of birdsong both online and internally in the Gentner lab, it is not necessary to use exact data from the paper to reproduce techniques or results, as the paper is designed to work beyond hyper specific vocalizations, but due to future research focusing on birds, I will still use birdsong. The first challenge for this project will be precise segmentation of birdsong into syllables. There are multiple computational methods, all of which have different benefits, it may be necessary to run multiple and segment by agreement between methods. 
From there, Sainburg 2020 extensively uses dimensionality reduction techniques, primarily UMAP, and develops methods for rating the dimensional space from the clustering that appears. A challenge here will be developing both methods of contrasting these results to a sparse space a la Smith & Lewicki, and/or fitting the sparse coder into the pipeline prior to UMAP to compare to Fig 3 in Sainburg 2020. Sainburg 2020 has many figures illustrating the technique on a variety of datasets/vocalizations. For my purposes, my plan is to restrain this to birdsong and human vocalizations to assess the clusterability differences between those with both different data and different models.  


Sainburg 2020 link: https://journals.plos.org/ploscompbiol/article?id=10.1371/journal.pcbi.1008228
Smith & Lewicki 2006 link: https://nature.com/articles/nature04485    



## Methods

### Power Analysis

Original effect size, power analysis for samples to achieve 80%, 90%, 95% power to detect that effect size.  Considerations of feasibility for selecting planned sample size.

### Planned Sample

Planned sample size and/or termination rule, sampling frame, known demographics if any, preselection rules if any.

### Materials

All materials - can quote directly from original article - just put the text in quotations and note that this was followed precisely.  Or, quote directly and just point out exceptions to what was described in the original article.

### Procedure	

Can quote directly from original article - just put the text in quotations and note that this was followed precisely.  Or, quote directly and just point out exceptions to what was described in the original article.

### Analysis Plan

Can also quote directly, though it is less often spelled out effectively for an analysis strategy section.  The key is to report an analysis strategy that is as close to the original - data cleaning rules, data exclusion rules, covariates, etc. - as possible.  

**Clarify key analysis of interest here**  You can also pre-specify additional analyses you plan to do.

### Differences from Original Study

Explicitly describe known differences in sample, setting, procedure, and analysis plan from original study.  The goal, of course, is to minimize those differences, but differences will inevitably occur.  Also, note whether such differences are anticipated to make a difference based on claims in the original article or subsequent published research on the conditions for obtaining the effect.

### Methods Addendum (Post Data Collection)

You can comment this section out prior to final report with data collection.

#### Actual Sample
  Sample size, demographics, data exclusions based on rules spelled out in analysis plan

#### Differences from pre-data collection methods plan
  Any differences from what was described as the original plan, or “none”.


## Results


### Data preparation

Data preparation following the analysis plan.
	
```{r include=F}
### Data Preparation

#### Load Relevant Libraries and Functions

#### Import data

#### Data exclusion / filtering

#### Prepare data for analysis - create columns etc.
```

### Confirmatory analysis

The analyses as specified in the analysis plan.  

*Side-by-side graph with original graph is ideal here*

### Exploratory analyses

Any follow-up analyses desired (not required).  

## Discussion

### Summary of Replication Attempt

Open the discussion section with a paragraph summarizing the primary result from the confirmatory analysis and the assessment of whether it replicated, partially replicated, or failed to replicate the original result.  

### Commentary

Add open-ended commentary (if any) reflecting (a) insights from follow-up exploratory analysis, (b) assessment of the meaning of the replication (or not) - e.g., for a failure to replicate, are the differences between original and present study ones that definitely, plausibly, or are unlikely to have been moderators of the result, and (c) discussion of any objections or challenges raised by the current and original authors about the replication attempt.  None of these need to be long.